\begin{abstract}
	The rapid expansion of 6G edge computing and intelligent services has introduced new challenges in federated orchestration due to massive device density, heterogeneous tasks, and strict latency constraints. Existing federated reinforcement learning methods lack semantic awareness in task placement and suffer from high communication cost and unstable convergence when scaled to large edge networks. These limitations lead to elevated latency, higher power consumption, reduced semantic fidelity, and inefficient communication, particularly under dynamic, large-scale workloads.

	To address these challenges, we propose FedSemGNN, a semantic-aware and hierarchical federated reinforcement learning framework that integrates graph neural network-based topology encoding with Proximal Policy Optimization (PPO) for edge service orchestration. FedSemGNN embeds service requests into 128-dimensional semantic spaces using Sentence-BERT (\texttt{all-MiniLM-L6-v2}) and employs a two-layer graph convolutional network to capture network topology and resource constraints. Edge servers are organized into 10 semantic-topological clusters with two-level hierarchical aggregation: intra-cluster synchronization every 10 epochs and inter-cluster global synchronization every 50 epochs. The lightweight actor-critic architecture with only 16 hidden units per layer enables ultra-low orchestration latency while maintaining high reward performance through prioritized experience replay and differential privacy mechanisms.
	
	We implement FedSemGNN in EdgeSimPy, a high-fidelity edge computing simulator, and evaluate it across 6 heterogeneous edge servers and 500 mobile users over 500 orchestration steps. Large-scale behavior up to 10,000 nodes is analyzed through validated mathematical extrapolation using power-law and logarithmic scaling models fitted to empirical base results. FedSemGNN is compared with seven competitive baselines including FlatFedPPO, HierFedPPO, HSQF, Random, GreedyPower, GreedyLatency, and FirstFit.
	
	Experimental results demonstrate that FedSemGNN achieves exceptional performance across all metrics: normalized reward of 0.91 (2.9$\times$ higher than Random and 1.9$\times$ higher than GreedyPower), ultra-low orchestration latency of 0.36 ms per step (315$\times$ faster than FlatFedPPO and 222$\times$ faster than HierFedPPO), energy consumption of 72.1 W (36.2$\times$ lower than FlatFedPPO and 14.7$\times$ lower than HierFedPPO), and communication efficiency of 1.394 reward per MB. FedSemGNN maintains 100\% semantic fidelity with zero service migrations compared to 6,840 migrations in FlatFedPPO and 1,181 in HierFedPPO. These results demonstrate that FedSemGNN is a scalable and efficient orchestration framework for federated learning in next-generation 6G edge intelligence.
\end{abstract}
