\section{Conclusion}
\label{sec:conclusion}

This paper presented \textbf{FedSemGNN}, a hierarchical, semantic-aware federated reinforcement learning framework designed for 6G edge orchestration. By jointly leveraging custom continual learning semantic embeddings in 16-dimensional spaces, two-layer GCN-based topology encoding with symmetric normalization, and hierarchical PPO training with dual-level aggregation ($K_1 = 10$ epochs local, $K_2 = 50$ epochs global), FedSemGNN achieves ultra-low latency (0.36 ms), exceptional energy efficiency (72.1 W), and perfect semantic fidelity (100\%) with zero service migrations.

Implemented within the EdgeSimPy simulator across a heterogeneous edge infrastructure with mobile users following realistic mobility patterns over 1,000 orchestration timesteps, and benchmarked against four competitive baselines (FlatFedPPO, HierFedPPO, HSQF, RandomPlacement), FedSemGNN demonstrated exceptional performance across all metrics. The framework achieves normalized reward of 0.91, representing 8.4\% improvement over FlatFedPPO and 16.7\% improvement over HierFedPPO. Orchestration latency is 315 times faster than FlatFedPPO (114.25 ms) and 222 times faster than HierFedPPO (80.71 ms). Power consumption is 36.2 times lower than FlatFedPPO (2607.9 W) and 14.7 times lower than HierFedPPO (1057.4 W). Communication efficiency reaches 1.394 reward per MB, which is 174 times better than FlatFedPPO.

The evaluation yielded four key insights:
\begin{enumerate}
	\item \textbf{Semantic embeddings enable perfect placement stability}: Custom continual learning semantic encoder with cosine similarity matching achieves 100\% semantic fidelity with zero migrations, compared to 6,840 migrations in FlatFedPPO and 1,181 in HierFedPPO.
	
	\item \textbf{GCN topology encoding provides spatial coherence}: Two-layer graph convolutional networks with cached normalization capture multi-hop dependencies across mesh-connected infrastructure, enabling spatially aware orchestration decisions.
	
	\item \textbf{Hierarchical federated learning enables extreme scalability}: Dual-level aggregation (local every 10 epochs, global every 50 epochs) reduces communication overhead by 161 times while maintaining convergence quality, with mathematical extrapolation demonstrating logarithmic scaling to 10,000 nodes and beyond.
	
	\item \textbf{Compact neural architectures enable real-time operation}: Lightweight actor-critic networks with only 16 hidden units per layer, combined with prioritized experience replay (buffer size 20,000) and differential privacy ($\sigma_{\text{DP}} = 1.0$), achieve sub-millisecond orchestration latency suitable for 6G control-plane requirements.
\end{enumerate}

The reward function $R_t = \max(0, 10{,}000 - (P_t + 0.01 \cdot L_t))$ with exponential moving average smoothing ($\alpha_{\text{smooth}} = 0.2$) successfully balances energy efficiency and response latency while implicitly encouraging semantic fidelity through state representation design.

Collectively, these findings establish FedSemGNN as a principled, empirically validated foundation for intelligent, privacy-preserving orchestration in next-generation 6G edge computing systems. The framework's combination of semantic awareness, graph-based spatial reasoning, hierarchical coordination, and compact neural architectures addresses critical challenges in scalability, latency, energy efficiency, and semantic coherence. This positions FedSemGNN as a viable solution for real-world edge-AI deployments in emerging 6G networks.
